<<<<<<< Updated upstream
---
title: "Lab 4"
author: "P-Hackers"
date: "9/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) 
```
## Over-Arching Goals

### 1) Analysis of premature births

A premature birth is defined as one that occurs before a gestational age of 270 days. We want to see the linear corilation between premature births from mothers that smoked durning their pregnancy vs. the mothers that did not smoke during their pregnancy.

### 2) Analysis of birth weights at gestational ages

To test the Surgeon General's assertion that newborns of mothers who smoke have lower birth weights at every gestational age (number of days into the pregnancy when the child is born), we want to create graphs to visualize the linear relationship between smoking mothers and the babies weight at the gestational age. 

<p>&nbsp;</p>



## Team Hypothesis
<p>&nbsp;</p>

### Question 1

We believe that there will be a strong positive corilation between premature births and mothers smoking while pregnant.

### Question 2

There will be a positive correlation between the length of gestation and birth weight. Because the baby develops during gestation, it follows that a longer period of gestation would result in higher birth weight. We also predict that the children of smoking mothers will weigh less than children of non-smoking mothers.  

<p>&nbsp;</p>
## Why Important

### Question 1

Determining the effect smoking has on the rate of pre-term birth is of vital importance to new mothers. If smoking harms the health of a newborn child, the public must be informed of these risks so that they may take the necessary steps to protect themselves and their families.  

### Question 2

<p>&nbsp;</p>
## More Detailed Why Important 

### Question 1

If it's found that smoking harms pregnant women and their babies, it's reasonable to assume that expecting mothers would stop, or severely limit their smoking. This would deal a catastrophic blow to cigarette manufacturers. Additionally, it would provide further information as to why babies are born prematurely and could open up new avenues of inquiry as to the effects of smoking on pregnancy and birth. 

### Question 2

<p>&nbsp;</p>


## Additional Team Study

### Smokers vs. NonSmokers

**Threats to the Valididty of the Experiment:**


It's imperative to note each of the variables taken into consideration during the experiment. While more variables result in less error, it's nearly impossible to consider all of the threats. Note that the variables considered while running the numbers are as follows: ID, birthday, gestation length, birth weight, mother's/father's race, mother's/father's age, mother's/father's education level, mother's/father's pre-pregnancy weight, marital status of the mother, family income, mother's smoking, how long the mother has been smoking/ how long since she quit smoking, the number of cigarettes the mother smokes per-day, and if the baby was born prematurely. 

<p>&nbsp;</p>

#### Here are some graphs showing the variables that we thought are simular between smoking mothers and non-smoking mothers:

```{r, warning=FALSE,echo=FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
ggplot(data=babies,mapping=aes(x=dht))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Fathers Height", x ="Fathers Height") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```


```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=mht))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Mothers Height", x ="Mothers Height") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```



```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=dage))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Fathers Age", x ="Fathers Age") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```



```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=mrace))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Mothers Race", x ="Mothers Race") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```



```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=mrace))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Fathers Race", x ="Fathers Race") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```



```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=date))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Babies Birthday", x ="Babies Birthday") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```


#### Here are some graphs showing the variables that we thought are noticably different between smoking mothers and non-smoking mothers:


```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=mpregwt))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Mothers Weight", x ="Mothers Weight") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```


```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=dwt))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Fathers Weight", x ="Fathers Weight") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```


```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=med))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Mothers Education", x ="Mothers Education") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```


```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=ded))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Fathers Education", x ="Fathers Education") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```


```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=bwtoz))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Babies Weight", x ="Babies Weight (IDS)") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```


```{r, warning=FALSE,echo=FALSE}
ggplot(data=babies,mapping=aes(x=gestation))+
  geom_bar()+
  facet_wrap(~smoke) +
  labs(title = "Gestation Period", x ="Gestation Period") +
  theme_classic() +
  theme(plot.title = element_text(color="black",face="bold"))

```




### Plot for Hypothesis 1

```{r,warning=FALSE, echo=FALSE}
Snames <- c(
  '0' = "Not-Premature",
  '1' = "Premature",
  'NA' = "Not Asked"
)

ggplot(data=babies,mapping=aes(x=bwtoz, group=smoke, color = smoke)) +
  geom_freqpoly() +
  facet_wrap(~Premature, labeller = as_labeller(Snames))+
  labs(title = "The Effect of a Premature Birth on Newborn Weight",x = "Weight of the New Born (Oz)")+
scale_color_continuous(label = c("0.00"= "Nonsmoker","0.25"="Once a Smoker,but not now","0.50"= "Smoker until Pregnancy", "0.75" = "Smoker everyonce and a while", "1.00"= "Full-Smoker")) +
  theme_classic()+
 theme( plot.title = element_text(color = "black", face = "bold"))
  
```

#### Conclusion
The data proved our first hypothesis to be correct. It showed that full-term births were mainly attributed to non-smokers and that premature births were primarily ascribed to smokers. What's more, the sample data was mainly made up of non-smokers and smokers. There weren't many people reporting as in-between. 

<p>&nbsp;</p>

### Plot for Hypothesis 2

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = babies, mapping = aes(x=bwtoz, y = gestation, color = Premature)) +
  geom_jitter() +
  geom_smooth(method = lm) +
  facet_wrap(~smoke) +
  labs(title="Gestation v. Birth Weight") +
  theme(plot.title = element_text(color="black",face="bold"))
```

```{r, echo=FALSE}
lm(gestation~bwtoz, data = babies)
```

#### Conclusion


We hypothesized that there would be a positive correlation between birth weight and gestation. After plotting and analyzing the data, we were proven correct. In our lm, the slope ( birth weight/gestation) is .3422 which indicates a positive correlation. Our secondary hypothesis concerning the effects of smoking on birth weight was slightly correct. We observed a weak negative correlation between birth weight and smoking. Our model showed that babies whose mothers smoke have a slightly lower birth weight than babies whose mothers do not. Lastly, the difference in color exhibited between premature and non-premature babies is interesting as it further demonstrates the correlation between gestation and birth weight. 

<p>&nbsp;</p>

### Recommendation 






## Individual Studies 
<p>&nbsp;</p>

### Julia Foley 
<p>&nbsp;</p>

#### Question

What is the correlation between a mother's pregnant weight and a baby's birth weight? Do babies with mothers who smoke have a higher or lower birth weight? Are babies whose mothers smoke more likely to be premature than babies whose mothers don't smoke? 

#### Hypothesis

I believe that there will be a positive correlation between birth weight and mother's pregnant weight. Additionally, I believe that Premature babies will have a lower birth weight than babies who are not premature. Lastly, I believe that in the premature section, there will be more babies whose mothers smoke than babies whose mothers did not smoke. 

#### Birth Weight v. Mother's Weight in Relation to Smoking

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = babies, mapping = aes(x=bwtoz, y = mpregwt, color = smoke)) +
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(~Premature)
```
```{r, echo = FALSE}
lm(mpregwt~bwtoz, data = babies)
```

#### Findings

I found that my hypothesis that there would be a positive correlation between a mother's pregnant weight and birth weight was correct. As shown by the lm slope, there is a weak positive correlation. Additionally, the plots show that there is a slight shift to the left in the premature graph as opposed to the not premature graph. This means that babies who are premature have a lower birth weight. Additionally, the data shows that there are more premature babies with mothers who smoked than with mothers who didn't smoke. This supports the two main questions because the plots show the relationship between a mother's weight, birth weight, premature, and smoking. Additionally, the data supports the two main questions because it shows that there is a negative correlation between premature and smoking, and a positive correlation between premature and smoking.  

<p>&nbsp;</p>
### Declan Franklin 
<p>&nbsp;</p>

#### Question: 

Does Income effect weather the mother smokes or not?

#### Hypothesis:

If both parents have a higher income, the mother is less likely to smoke during pregnancy.

```{r,echo=FALSE, warning=FALSE}

ggplot(data=babies,mapping=aes(x=inc, group=smoke, fill = smoke)) +
  geom_density(alpha=0.4) +
  scale_x_continuous(labels=c("0.0"="< $2500","2.5"="~$6250","5.0"="~$12500","7.5"="~$18750","9.0"= "~$22500")) +
  labs(title = "The Income of Smokers vs Non-Smokers", caption = "Kernel Density Graph", x="Family Income")+
  scale_fill_continuous(label = c("0.00"= "Nonsmoker","0.25"="Once a Smoker,but not now","0.50"= "Smoker until Pregnancy", "0.75" = "Smoker everyonce and a while", "1.00"= "Full-Smoker"))+
  theme_classic()+
 theme( plot.title = element_text(color = "black", face = "bold"))
  
```

#### Findings:

I found that most new families between $2500 - $7000. I also found that the non-smokers make more money than the smokers. This info could give insight to our team’s data by giving us insight on what the families can bring to the table while pregnant.

<p>&nbsp;</p>
### Nisia Pinto
<p>&nbsp;</p>

**Question:**

Mother Education Vs Dad Education and Parity:
Do couples with higher  education tend to have less kids than the ones with lower education?**

**Hypothesis:**

Couples with higher educational levels have less kids than the ones with lower educational levels.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = babies , mapping = aes(x=ded, y = med, color=parity)) +
  geom_point() +
  scale_color_gradient(low="blue", high="orange")+
  geom_jitter() +
  geom_smooth(method = lm) + facet_wrap(~parity) +
  theme(plot.title=element_text(color="orange",face="bold", hjust=0.5))+
  labs(title="Mother Education Vs Father Education in relation to Parity")
```
```{r echo=FALSE}
babies <- na.omit(babies)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results=FALSE }
cor(babies$med,babies$ded)
cor(babies$med,babies$parity)
cor(babies$ded,babies$parity)
```

**Findings:**

Correlations:
**Med vs Ded:** 0.5992664    **Med vs Parity** : -0.1760948    **Ded vs Parity:** -0.1312375

The mather and father's education are positively correlated. However, parity is negatively correlated to both of them. I was able to answer my question and confirm my hypothesis, people with higher education have less babies than less educated people. This can be due to the addition of other different factors:college tuitions, retirement money, health care, or salary.
In addition, we can see that the positive correlations are moderate and the negative are weak.


<p>&nbsp;</p>
### Andres Acevedo
<p>&nbsp;</p>

### Question:
1) Given the dataset we have, is there enough information to make a conclusion as to whether smoking is the main factor in decreased birth weights? Or could lower birthweights be explained better by a combination of factors including smoking as well as the height and weight of the mother and father?

2) Are there really that many more premature births associated with smokers as opposed to non-smokers? What else could cause a premature birth?

#### Hypothesis
There will be an insignificant difference in the number of premature births attributed to smoking mothers and those attributed to non-smoking mothers.

```{r,echo=FALSE, warning=FALSE}
ggplot(data = babies) +
  geom_bar(mapping = aes(x = Premature), fill = 'steelblue') +
  facet_grid(~smoke) + 
  theme(strip.text.x = element_text(size = 12, color = "steelblue", face = "bold.italic" )) +
  ggtitle('Number of Premature Births: Non-Smoking vs Smoking vs N/A')
```

#### Findings

I found that my hypothesis was correct. After plotting the number of premature births attributed to smokers vs the number attributed to non-smokers, I found an insignificant difference between the number of premature births reported by smokers relative to the number reported by non-smokers. This could be because of the presence of several confounding variables that could help explain the gestational period and birth weights of children. Weight and height of the parents, as well as income and education, can play a role in the physical and emotional health of the mother in the same way that smoking can.


<p>&nbsp;</p>
## What each Member did 
<p>&nbsp;</p>

### Julia Foley


#### Individual 

I created a plot using ggplot, geom_point, geom_smooth, lm, facet_wrap, and color = smoke. I also used lm to find the equation of the line. This helped me to compare the relationship between mother's pregnant weight, birth weight, premature, and smoking. Additionally, I made a question and a hypothesis, and I reported my findings.

#### Team 

I created the document, I wrote the hypothesis and the conclusion for question 2. I also helped out with plot 2. 

<p>&nbsp;</p>
### Declan Franklin 

#### Individual

I created my own study on how income differs from smokers and non-smokers. I used geom_density and a facet_wrap to visualize the data. I also used different labelling techniques so that the plots formatted nicely.

#### Team 

For the team: I created all the plots of potentially similar and different variables, I answered the Validity question, I formatted the HTML for aesthetic purposes, and I did the final review of everyone’s Code to make sure it ran correctly and neatly.

<p>&nbsp;</p>
### Andres Acevedo

#### Individual

I created a multitude of plots relating variables in the dataset 'babies'. I noticed there were very few variables that were even moderately correlated save for gestation vs baby weight. I used a geom_bar() object to count the number of premature births for smoking mothers as opposed to non-smoking mothers. I used a facet_wrap() to compare the smokers to the non-smokers. I applied a theme to the labels across the top of the plot. I filled in the bars of the graph with the color 'steelblue' and I applied a title to the graph using ggtitle().

#### Team

I went through all of the team language in our Markdown file and touched it up so that it would be more suitable to a formal setting like a meeting with a client. I formalized the language and corrected any mistakes or typos that I found.